---
title: "PA1_template"
author: "Ali CHAOUCHE"
date: "5/13/2020"
output:
  word_document: default
  html_document: default
---
output: 
  html_document:
    keep_md: true
    
rmarkdown::render(  input = 'file.RMD' , output_file = 'file.md', output_format=   md_document( ) )  
```{r, echo = TRUE,include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# **Reproducible Research Coursera**
# **Peer-graded Assignment Course Project 1**

```{r results='hide', echo = TRUE,message=FALSE, warning=FALSE}
# Install R libraries
library(data.table)
library(dplyr)
library(knitr)
library(Hmisc)
```


## **QUESTION 1: Code for reading in the dataset and/or processing the data** 
Load the data (i.e.read.csv())
Process/transform the data (if necessary) into a format suitable for your analysis
```{r echo = TRUE,}
# Read data
activity <- read.csv("activity.csv", header=TRUE)
str(activity)
dim(activity)
head(activity,5)
tail(activity,5)
```
## **QUESTION 2: Histogram of the total number of steps taken each day**
Calculate the total number of steps taken per day
If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day
Calculate and report the mean and median of the total number of steps taken per day

```{r echo = TRUE,}
# Calculate the total number of steps taken per day
totalsteps <- with(activity, tapply(activity$steps, as.factor(activity$date), sum, na.rm = T))
# plot histogram
hist(totalsteps,
     main="Histogram of the total number of steps taken each day",
     xlab="Total number of steps",
     ylim=c(0,20),
     col="grey", breaks=10,
     freq=TRUE)
```

## **QUESTION 3: Mean and median number of steps taken each day**
```{r echo = TRUE}
 mean (totalsteps)
 median (totalsteps)
```
## **QUESTION 4: Time series plot of the average number of steps taken**
Make a time series plot (i.e. type = "l" of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
``` {r echo = TRUE}
timeseries <- aggregate(activity$steps, by=list(activity$interval), FUN=mean, na.rm=TRUE)

names(timeseries)<-c("interval","mean")
 
plot(timeseries,
main="Time series plot of the average number of steps taken",
xlab="Interval", xlim=c(0,2500),
ylab="Average steps ",
type="l",
col="black")
 
```

## **QUESTION 5: The 5-minute interval that, on average, contains the maximum number of steps**

 ```{r echo = TRUE}
 maxnumberofsteps<- which.max(timeseries$mean)
 intervalmax <- timeseries[maxnumberofsteps,]
 print(intervalmax)
 ```
 
## **QUESTION 6 : Code to describe and show a strategy for imputing missing data**
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with 
```{r echo = TRUE}
 missing_data <- which(is.na(activity$steps))
```
Total missing values is equal to: 

```{r echo = TRUE}
length(missing_data)
```
Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r echo = TRUE}
imputedata<-activity

imputedata$steps <- impute(activity$steps, fun=mean)
```
Total number of steps taken each day after missing values are imputed
```{r echo = TRUE}
totalimpute <- with(imputedata, tapply(imputedata$steps, as.factor(imputedata$date), sum, na.rm = T))
```
## **QUESTION 7 : Histogram of the total number of steps taken each day after missing values are imputed** 

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r echo = TRUE}
hist(totalimpute,
     main="Histogram of the total number of steps taken each day (imputed)",
     xlab="Total number of steps",
     ylim=c(0,25),breaks=10,
     col="grey",
     freq=TRUE)
```


Do these values differ from the estimates from the first part of the assignment ? 
What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r echo = TRUE}
summary(totalimpute)
```

Imputing the missing datas gives differences in mean, median values and histograms behaviour.  

## **QUESTION 8 : Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends**
Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r echo = TRUE}
imputedata['type_of_day'] <- weekdays(as.Date(imputedata$date))
imputedata$type_of_day[imputedata$type_of_day  %in% c('Saturday','Sunday') ] <- "weekend"
imputedata$type_of_day[imputedata$type_of_day != "weekend"] <- "weekday"
# Convert type_of_day from character to factor
imputedata$type_of_day <- as.factor(imputedata$type_of_day)
# Calculate average steps by interval across all days
finaldata <- aggregate(steps ~ interval + type_of_day, imputedata, mean)
```
Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r results='hide', echo = TRUE, message=FALSE, warning=FALSE}

qplot(interval, 
      steps, 
      data = finaldata, 
      type = 'l', 
      geom=c("line"),
      xlab = "Interval", 
      ylab = "Number of steps", 
      main = "") +
        facet_wrap(~ type_of_day, ncol = 1)
```
